---
title: "Untitled"
output: html_document
date: '2023-07-10'
---
```{r}
setwd("/Users/samd/OneDrive\ -\ Northwestern\ University/Primate/Scripts")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Doing deseq2 on primate skin data ..first on subset of data with only antibiotics and lipid synthesis

```{r}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("DESeq2")
```

```{r}
setwd("/Users/samd/OneDrive\ -\ Northwestern\ University/Primate/Scripts")
library(DESeq2)
library(ggplot2)
```
```{r}

cts <- read.table("deseqOTUtable.txt", sep='\t',header=TRUE)
head(cts)

ctsAgg=aggregate(cts[ , 2:76], by = list(cts$OTUID), FUN = sum)                # Consolidate duplicates
head(ctsAgg)  
dim(cts)
dim(ctsAgg)

ctsMat=as.matrix(ctsAgg)
head(ctsMat)
ctsMatm <- ctsMat[,-1]
rownames(ctsMatm) <- ctsMat[,1]
head(ctsMatm)


f <- file.choose("metadataDeseq.txt")
meta <- read.table(f, sep="\t",header=TRUE,row.names=1)
head(meta)

dds <- DESeqDataSetFromMatrix(countData = ctsMatm,
                              colData = meta,
                              design= ~ type)
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients
res <- results(dds, name="condition_trt_vs_untrt")
# or to shrink log fold changes association with condition:
res <- lfcShrink(dds, coef="condition_trt_vs_untrt", type="apeglm")

all(rownames(meta) == colnames(ctsMatm))
rownames(ctsMatm) <- make.names(names_countdata, unique=TRUE)

ctsMatm <- ctsMat[,-1]
names_count <- ctsMat[,1]
rownames(ctsMatm) <- make.names(names_count, unique=TRUE)

ctsMatm <- noquote(ctsMatm)
ctsMatmI=as.integer(ctsMatm)
dds <- DESeqDataSetFromMatrix(countData = ctsMatmI,
                              colData = meta,
                              design= ~ type)
 #doesnt work
 
 class(ctsMatm)
 str(ctsMatm)

 #chooing deseqOTUtable.csv manually with this file.choose code because it isnt working
 f2=file.choose("deseqOTUtable.csv")
cts2 <- read.csv(f2,header=TRUE)
head(cts2)

cts2Agg=aggregate(cts[ , 2:10], by = list(cts2$OTUID), FUN = sum) 
  
cts3 <- cts2Agg[,-1]
names_cts <- cts2Agg[,1]
rownames(cts3) <- make.names(names_cts, unique=TRUE)
head(cts3)

dds <- DESeqDataSetFromMatrix(countData = cts3,
                              colData = meta,
                              design= ~ type)
                              
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients
res <- results(dds, name="type_human_vs_chimp")
# or to shrink log fold changes association with condition:
res <- lfcShrink(dds, coef="condition_trt_vs_untrt", type="apeglm")

humanVg=results(dds, contrast=c("type","human","gorilla"))
res2=humanVg[order(-humanVg$padj), "Number", drop = FALSE]

humanVg2=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(humanVg2), file="deseq2resultsHumanVg2.csv")


ctsG <- read.csv("DeseqGenusTable.csv", sep=',',header=TRUE)
head(ctsG)

#cts2Agg=aggregate(cts[ , 2:10], by = list(cts2$OTUID), FUN = sum) 
  
ctsG2 <- ctsG[,-1]
names_ctsG <- ctsG[,1]
rownames(ctsG2) <- make.names(names_ctsG, unique=TRUE)
head(ctsG2)

meta <- read.table("metadataDeseq.txt", sep="\t",header=TRUE,row.names=1)
meta

ddsG <- DESeqDataSetFromMatrix(countData = ctsG2,
                              colData = meta,
                              design= ~ type)
                              
ddsG <- DESeq(ddsG)

humanVgG=results(ddsG, contrast=c("type","human","gorilla"))
humanVgG2=humanVgG[order(humanVgG$padj),]
write.csv(as.data.frame(humanVgG2), file="deseq2resultsHumanVgG2.csv")

humanVchimpG=results(ddsG, contrast=c("type","human","chimp"))
humanVchimpG2=humanVchimpG[order(humanVchimpG$padj),]
write.csv(as.data.frame(humanVchimpG2), file="deseq2resultsHumanVchimpG2.csv")

chimpVgorG=results(ddsG, contrast=c("type","chimp","gorilla"))
chimpVgorG2=chimpVgorG[order(chimpVgorG$padj),]
write.csv(as.data.frame(chimpVgorG2), file="deseq2resultsChimpVgorG2.csv")

```

Now trying pathway independet of taxa Deseq..i switched directory at top to Scripts

```{r}
meta <- read.table("Primate_metadata.txt", sep="\t",header=TRUE,row.names=1)
meta
```
```{r}
cts2 <- read.csv("pathway_type.csv",header=TRUE)
head(cts2)

#cts2Agg=aggregate(cts[ , 2:10], by = list(cts2$OTUID), FUN = sum)   no need to Agg for this
  
cts3 <- cts2[,-1]
names_cts <- cts2[,1]
rownames(cts3) <- make.names(names_cts, unique=TRUE)
head(cts3)

#match rows to meta columns

meta2 <- meta[row.names(meta) %in% colnames(cts2),]
meta2
meta3=meta2[match(colnames(cts2[,-1]), rownames(meta2)),]
meta3

dds <- DESeqDataSetFromMatrix(countData = cts3,
                              colData = meta3,
                              design= ~ Species)
                              
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients
res <- results(dds, name="Species_human_vs_Gorilla.gorilla.gorilla")
# or to shrink log fold changes association with condition:
res <- lfcShrink(dds, coef="condition_trt_vs_untrt", type="apeglm")

humanVg=results(dds, contrast=c("Species","human","Gorilla_gorilla_gorilla"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

humanVg2=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(humanVg2), file="deseq2PathwayIndresultsHumanVg.csv")
```
now trying chimp..i think need to change order of names
```{r}
humanVc=results(dds, contrast=c("Species","human","Pan_troglodytes"))
res2=humanVg[order(-humanVc$padj), "pvalue", drop = FALSE]

humanVc2=humanVc[order(humanVg$padj),]

write.csv(as.data.frame(humanVc2), file="deseq2PathwayIndresultsHumanVchimp.csv")
```

Now trying keggs

```{r}
cts2 <- read.csv("kegg2koname.csv",header=TRUE,row.names=1)
head(cts2)

dds <- DESeqDataSetFromMatrix(countData = cts2,
                              colData = meta3,
                              design= ~ Species)
                              
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

humanVg=results(dds, contrast=c("Species","human","Gorilla_gorilla_gorilla"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

humanVg2=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(humanVg2), file="deseq2PathwayIndresultsHumanVgKegg.csv")
```

```{r}
humanVg=results(dds, contrast=c("Species","human","Pan_troglodytes"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

humanVg2=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(humanVg2), file="deseq2PathwayIndresultsHumanVchimpKegg.csv")
```

```{r}
cts2 <- read.csv("keggpath2class.csv",header=TRUE,row.names=1)
head(cts2)

dds <- DESeqDataSetFromMatrix(countData = cts2,
                              colData = meta3,
                              design= ~ Species)
                              
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

humanVg=results(dds, contrast=c("Species","human","Gorilla_gorilla_gorilla"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

humanVg2=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(humanVg2), file="deseq2PathwayIndresultsHumanVgClass.csv")
```
```{r}

humanVg=results(dds, contrast=c("Species","human","Pan_troglodytes"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

humanVg2=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(humanVg2), file="deseq2PathwayIndresultsHumanVchimpClass.csv")
```

Now doing by species level

```{r}
meta <- read.table("pathwayspecies_metadata.txt", sep="\t",header=TRUE,row.names=1)
```

```{r}
cts2 <- read.csv("pathwaytype_species_Anti.csv",header=TRUE,row.names=1)
head(cts2)

cts2= as.matrix(cts2+1)

dds <- DESeqDataSetFromMatrix(countData = cts2,
                              colData = meta,
                              design= ~ Human)

                              
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

humanVg=results(dds, contrast=c("Human","yes","no"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

humanVg2=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(humanVg2), file="deseq2PathwaySpeicesresultsHumanVnon.csv")
```

Now trying three factors / 3 species so cts2 is staying same as above

```{r}
dds <- DESeqDataSetFromMatrix(countData = cts2,
                              colData = meta,
                              design= ~ Species)

                              
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

humanVg=results(dds, contrast=c("Species","human","chimp","gorilla"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

humanVg2=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(humanVg2), file="deseq2PathwaySpeicesresultsHumanVchimpVgor.csv")
```
Seems like it doesnt work and only accepts 2

plot code for path3ay species of human vs non human Anit subset

```{r}
ars=read.table("deseq2PathwaySpeicesresultsHumanVnon.csv",sep=',',row.names=1,header=TRUE)
ars
```
#Might need to redo pathways below this one again? not sure
```{r}
ars2=subset(ars,padj<0.1)
OTUs=rownames(ars2)

pARS=ggplot(ars2,aes(reorder(OTUs,-log2FoldChange),log2FoldChange,fill=log2FoldChange < 0))+
    geom_bar(stat="identity")+
    geom_errorbar(aes(ymin=log2FoldChange-lfcSE, ymax=log2FoldChange+lfcSE), width=.2, alpha=0.6,
                 position=position_dodge(.9)) +
#ylim(-7,5.5)+
    theme_classic()+
    ylab("log2FoldChange")+
    labs(title = "Antibiotic resistant microbes in human and non-human primates",fill="") +
  scale_fill_manual(labels = c("Human", "Non-human"), values = c("#ff6e54","#444e86" ))+
    theme(axis.text.x = element_text(angle = 0,hjust=0.95,vjust=0.2,size=12),
              axis.text.y = element_text(size=12),
          axis.title.y = element_blank(),
          axis.title.x=element_text(size=12),
          title=element_text(size=12))+
   theme(plot.title = element_text(hjust = 0.65,vjust=1))+
    coord_flip()+
 
  theme(legend.text=element_text(size=12))
pARS
```


now doing pathway plot
```{r}
cts2 <- read.csv("pathway_type.csv",header=TRUE)
head(cts2)

#cts2Agg=aggregate(cts[ , 2:10], by = list(cts2$OTUID), FUN = sum)   no need to Agg for this
  
cts3 <- cts2[,-1]
names_cts <- cts2[,1]
rownames(cts3) <- make.names(names_cts, unique=TRUE)
head(cts3)

#match rows to meta columns

meta2 <- meta[row.names(meta) %in% colnames(cts2),]
meta2
meta3=meta2[match(colnames(cts2[,-1]), rownames(meta2)),]
meta3

dds <- DESeqDataSetFromMatrix(countData = cts3,
                              colData = meta3,
                              design= ~ human)
                              
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients


humanVg=results(dds, contrast=c("human","yes","no"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

humanVg2=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(humanVg2), file="deseq2PathwayIndresultsHumanVnon.csv")
```
```{r}
pw=read.table("deseq2PathwayIndresultsHumanVnon.csv",sep=',',row.names=1,header=TRUE)
pw
```


```{r}
pw2=subset(pw,padj<0.01)
OTUs=rownames(pw2)

pW=ggplot(pw2,aes(reorder(OTUs,-log2FoldChange),log2FoldChange,fill=log2FoldChange < 0))+
    geom_bar(stat="identity")+
    geom_errorbar(aes(ymin=log2FoldChange-lfcSE, ymax=log2FoldChange+lfcSE), width=.2, alpha=0.6,
                 position=position_dodge(.9)) +
ylim(-4.75,3)+
    theme_classic()+
    ylab("log2FoldChange")+
    labs(title = "Functional pathways in human and non-human primate skin microbiomes",fill="") +
  scale_fill_manual(labels = c("Human", "Non-human"), values = c("#ff6e54","#444e86" ))+
    theme(axis.text.x = element_text(angle = 0,hjust=0.95,vjust=0.2,size=12),
              axis.text.y = element_text(size=12),
          axis.title.y = element_blank(),
          axis.title.x=element_text(size=12),
          title=element_text(size=12))+
   theme(plot.title = element_text(hjust = 0.65,vjust=1))+
    coord_flip()+
 
  theme(legend.text=element_text(size=12))
pW


```


```{r}
pts=read.table("pathway_type_species.txt",header=TRUE,sep="\t")
pts
```
```{r}
smeta <- read.table("pathwayspecies_metadata.txt", sep="\t",header=TRUE,row.names=1)
#note that ptsCholine is the df i made below..
smeta2 <- smeta[row.names(smeta) %in% colnames(ptsCholine),]
smeta2
smeta3=smeta2[match(colnames(ptsCholine[,]), rownames(smeta2)),]
smeta3
```
```{r}
#ptsCHoline I made below
dds <- DESeqDataSetFromMatrix(countData = ptsCholine,
                              colData = smeta3,
                              design= ~ Human)
                              
dds <- DESeq(dds,fitType="mean")
resultsNames(dds) # lists the coefficients
```

```{r}
humanVg=results(dds, contrast=c("Species","human","Gorilla_gorilla_gorilla"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

humanVg2=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(humanVg2), file="deseq2PathwayIndresultsHumanVgClass.csv")
```

Trying Choline:

```{r}

pts=read.table("pathway_type_species.txt",header=TRUE,sep="\t")
# Subset the dataframe using dplyr
ptsCholine <- pts %>%
  filter(Pathway == "Choline-Degradation") %>%
  select(-Pathway) %>%
  tibble::column_to_rownames(var="featureID")
ptsCholine= as.matrix(ptsCholine+1)

dds <- DESeqDataSetFromMatrix(countData = ptsCholine,
                              colData = smeta3,
                              design= ~ Human)
                              
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

humanVg=results(dds, contrast=c("Human","yes","no"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

humanVg2=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(humanVg2), file="deseq2PathwayCholineSpeciesHumanVnon.csv")

~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  
pwc=read.table("deseq2PathwayCholineSpeciesHumanVnon.csv",sep=',',row.names=1,header=TRUE)
pwc
pwc2=subset(pwc,padj<0.05)
OTUs=rownames(pwc2)

pC=ggplot(pwc2,aes(reorder(OTUs,-log2FoldChange),log2FoldChange,fill=log2FoldChange < 0))+
    geom_bar(stat="identity")+
    geom_errorbar(aes(ymin=log2FoldChange-lfcSE, ymax=log2FoldChange+lfcSE), width=.2, alpha=0.6,
                 position=position_dodge(.9)) +
ylim(-4.75,3)+
    theme_classic()+
    ylab("log2FoldChange")+
    labs(title = "Choline degradation microbes in human and non-human primate skin microbiomes",fill="") +
  scale_fill_manual(labels = c("Human", "Non-human"), values = c("#ff6e54","#444e86" ))+
    theme(axis.text.x = element_text(angle = 0,hjust=0.95,vjust=0.2,size=12),
              axis.text.y = element_text(size=12),
          axis.title.y = element_blank(),
          axis.title.x=element_text(size=12),
          title=element_text(size=12))+
   theme(plot.title = element_text(hjust = 0.65,vjust=1))+
    coord_flip()+
 
  theme(legend.text=element_text(size=12))
pC




```
DOesnt find any sig

trying Antibiotic Biosynth
```{r}
pts=read.table("pathway_type_species.txt",header=TRUE,sep="\t")
# Subset the dataframe using dplyr
ptsCholine <- pts %>%
  filter(Pathway == "Antibiotic-Biosynthesis") %>%
  select(-Pathway) %>%
  tibble::column_to_rownames(var="featureID")
ptsCholine= as.matrix(ptsCholine+1)

dds <- DESeqDataSetFromMatrix(countData = ptsCholine,
                              colData = smeta3,
                              design= ~ Human)
                              
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

humanVg=results(dds, contrast=c("Human","yes","no"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

humanVg2=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(humanVg2), file="deseq2PathwayAntiSynthSpeciesHumanVnon.csv")

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  
pwc=read.table("deseq2PathwayAntiSynthSpeciesHumanVnon.csv",sep=',',row.names=1,header=TRUE)
pwc
pwc2=subset(pwc,padj<0.05)
OTUs=rownames(pwc2)

pAS=ggplot(pwc2,aes(reorder(OTUs,-log2FoldChange),log2FoldChange,fill=log2FoldChange < 0))+
    geom_bar(stat="identity")+
    geom_errorbar(aes(ymin=log2FoldChange-lfcSE, ymax=log2FoldChange+lfcSE), width=.2, alpha=0.6,
                 position=position_dodge(.9)) +
ylim(-4.75,3)+
    theme_classic()+
    ylab("log2FoldChange")+
    labs(title = "Choline degradation microbes in human and non-human primate skin microbiomes",fill="") +
  scale_fill_manual(labels = c("Human", "Non-human"), values = c("#ff6e54","#444e86" ))+
    theme(axis.text.x = element_text(angle = 0,hjust=0.95,vjust=0.2,size=12),
              axis.text.y = element_text(size=12),
          axis.title.y = element_blank(),
          axis.title.x=element_text(size=12),
          title=element_text(size=12))+
   theme(plot.title = element_text(hjust = 0.65,vjust=1))+
    coord_flip()+
 
  theme(legend.text=element_text(size=12))
pAS

```

```{r}
pts=read.table("pathway_type_species.txt",header=TRUE,sep="\t")
# Subset the dataframe using dplyr
ptsCholine <- pts %>%
  filter(Pathway == "Peptidoglycan-Biosynthesis") %>%
  select(-Pathway) %>%
  tibble::column_to_rownames(var="featureID")
ptsCholine= as.matrix(ptsCholine+1)

dds <- DESeqDataSetFromMatrix(countData = ptsCholine,
                              colData = smeta3,
                              design= ~ Human)
                              
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

humanVg=results(dds, contrast=c("Human","yes","no"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

humanVg2=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(humanVg2), file="deseq2PathwayPepSynthSpeciesHumanVnon.csv")

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  
pwp=read.table("deseq2PathwayPepSynthSpeciesHumanVnon.csv",sep=',',row.names=1,header=TRUE)
pwp
pwp2=subset(pwp,padj<0.01)
OTUs=rownames(pwp2)

pPS=ggplot(pwp2,aes(reorder(OTUs,-log2FoldChange),log2FoldChange,fill=log2FoldChange < 0))+
    geom_bar(stat="identity")+
    geom_errorbar(aes(ymin=log2FoldChange-lfcSE, ymax=log2FoldChange+lfcSE), width=.2, alpha=0.6,
                 position=position_dodge(.9)) +
#ylim(-4.75,3)+
    theme_classic()+
    ylab("log2FoldChange")+
    labs(title = "Peptidoglycan synth microbes in human and non-human primate skin microbiomes",fill="") +
  scale_fill_manual(labels = c("Human", "Non-human"), values = c("#ff6e54","#444e86" ))+
    theme(axis.text.x = element_text(angle = 0,hjust=0.95,vjust=0.2,size=12),
              axis.text.y = element_text(size=12),
          axis.title.y = element_blank(),
          axis.title.x=element_text(size=12),
          title=element_text(size=12))+
   theme(plot.title = element_text(hjust = 0.65,vjust=1))+
    coord_flip()+
 
  theme(legend.text=element_text(size=12))
pPS
```
```{r}
kp=read_tsv("kegg2koname.tsv")
kp=kp %>% column_to_rownames(var="OTUID")

```
```{r}
kmeta=read.table("Primate_metadata.txt", sep="\t",header=TRUE,row.names=1)
kmeta
```
```{r}
kmeta2 <- kmeta[row.names(kmeta) %in% colnames(kp),]
kmeta2
kmeta3=kmeta2[match(colnames(kp[,]), rownames(kmeta2)),]
kmeta3
```

```{r}
dds <- DESeqDataSetFromMatrix(countData = kp,
                              colData = kmeta3,
                              design= ~ human)
                              
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

humanVg=results(dds, contrast=c("human","yes","no"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

humanVg2=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(humanVg2), file="deseq2keggHumanVnon.csv")

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  
pwk=read.table("deseq2keggHumanVnon.csv",sep=',',row.names=1,header=TRUE)
pwk
pwk2=subset(pwk,padj<0.0000000000001)
OTUs=rownames(pwk2)

pK=ggplot(pwk2,aes(reorder(OTUs,-log2FoldChange),log2FoldChange,fill=log2FoldChange < 0))+
    geom_bar(stat="identity")+
    geom_errorbar(aes(ymin=log2FoldChange-lfcSE, ymax=log2FoldChange+lfcSE), width=.2, alpha=0.6,
                 position=position_dodge(.9)) +
#ylim(-4.75,3)+
    theme_classic()+
    ylab("log2FoldChange")+
    labs(title = "Kegg pathways in human and non-human primate skin microbiomes",fill="") +
  scale_fill_manual(labels = c("Human", "Non-human"), values = c("#ff6e54","#444e86" ))+
    theme(axis.text.x = element_text(angle = 0,hjust=0.95,vjust=0.2,size=12),
              axis.text.y = element_text(size=12),
          axis.title.y = element_blank(),
          axis.title.x=element_text(size=12),
          title=element_text(size=12))+
   theme(plot.title = element_text(hjust = 1,vjust=1))+
    coord_flip()+
 
  theme(legend.text=element_text(size=12))
pK
```
```{r}
counts = matrix(rnbinom(1000,mu=50,size=1),100,10)

```
redoing graph to make it look simpler  with new file deseq2keggHumanVnon_modified_forGraph
```{r}
pwk=read.table("deseq2keggHumanVnon_modified_forGraph.csv",sep=',',row.names=1,header=TRUE)
pwk
pwk2=subset(pwk,padj<0.0000000000001)
OTUs=rownames(pwk2)

pK=ggplot(pwk2,aes(reorder(OTUs,-log2FoldChange),log2FoldChange,fill=log2FoldChange < 0))+
    geom_bar(stat="identity")+
    geom_errorbar(aes(ymin=log2FoldChange-lfcSE, ymax=log2FoldChange+lfcSE), width=.2, alpha=0.6,
                 position=position_dodge(.9)) +
#ylim(-4.75,3)+
    theme_classic()+
    ylab("log2FoldChange")+
    labs(title = "Kegg pathways in human and non-human primate skin microbiomes",fill="") +
  scale_fill_manual(labels = c("Human", "Non-human"), values = c("#ff6e54","#444e86" ))+
    theme(axis.text.x = element_text(angle = 0,hjust=0.75,vjust=0.2,size=12),
              axis.text.y = element_text(size=12),
          axis.title.y = element_blank(),
          axis.title.x=element_text(size=12),
          title=element_text(size=12))+
   theme(plot.title = element_text(hjust = 0.75,vjust=1))+
    coord_flip()+
 
  theme(legend.text=element_text(size=12))
pK
```
pW, pARS, pK
building a cowplot

```{r}
install.packages("cowplot")
library(cowplot)
```
```{r}
plot_grid(pW)
```
~~~~Doing Face region analysis now have to use original labels instead of the merged version that lumped facial regions together
```{r}
cts <- read.csv("pathway_type.csv",header=TRUE,row.names=1)
cts
```
```{r}
meta=read.table("Primate_metadata.txt", sep='\t',header=TRUE,row.names=1)
meta
```
```{r}
meta <- subset(meta, !is.na(Location))
meta=subset(meta,Location!='human')

icol=colnames(cts)
irow=rownames(meta)

int=intersect(irow,icol)
cts2=cts[ ,int]
meta2=meta[int, ]
```
```{r}
dds <- DESeqDataSetFromMatrix(countData = cts2,
                              colData = meta2,
                              design= ~ Location)
                              
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

humanVg=results(dds, contrast=c("Location","Chin","Cheek"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

ChinVCheek=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(ChinVCheek), file="deseq2keggChinVCheek.csv")
```
Now Chin V Forehead

```{r}
humanVg=results(dds, contrast=c("Location","Chin","Forehead"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

ChinForehead=humanVg[order(humanVg$padj),]
ChinForehead

write.csv(as.data.frame(ChinForehead), file="deseq2keggChinVForehead.csv")
```
CheekForehead
```{r}
humanVg=results(dds, contrast=c("Location","Cheek","Forehead"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

CheekVForehead=humanVg[order(humanVg$padj),]
CheekVForehead

write.csv(as.data.frame(CheekVForehead), file="deseq2keggCheekVForeead.csv")
```
```{r}
ChinVCheek
pwd=data.frame(ChinVCheek)
```
```{r}

pwd2=pwd[1:10,]
OTUs=rownames(pwd2)

pK=ggplot(pwd2,aes(reorder(OTUs,-log2FoldChange),log2FoldChange,fill=log2FoldChange < 0))+
    geom_bar(stat="identity")+
    geom_errorbar(aes(ymin=log2FoldChange-lfcSE, ymax=log2FoldChange+lfcSE), width=.2, alpha=0.6,
                 position=position_dodge(.9)) +
#ylim(-4.75,3)+
    theme_classic()+
    ylab("log2FoldChange")+
    labs(title = "Kegg pathways in human and non-human primate skin microbiomes",fill="") +
  scale_fill_manual(labels = c("Chin", "Cheek"), values = c("#ff6e54","#444e86" ))+
    theme(axis.text.x = element_text(angle = 0,hjust=0.75,vjust=0.2,size=12),
              axis.text.y = element_text(size=12),
          axis.title.y = element_blank(),
          axis.title.x=element_text(size=12),
          title=element_text(size=12))+
   theme(plot.title = element_text(hjust = 0.75,vjust=1))+
    coord_flip()+
 
  theme(legend.text=element_text(size=12))
pK
```

Now ChinForehead

```{r}
pwd=data.frame(ChinForehead)
pwd2=pwd[1:10,]
OTUs=rownames(pwd2)

pK=ggplot(pwd2,aes(reorder(OTUs,-log2FoldChange),log2FoldChange,fill=log2FoldChange < 0))+
    geom_bar(stat="identity")+
    geom_errorbar(aes(ymin=log2FoldChange-lfcSE, ymax=log2FoldChange+(lfcSE)), width=.2, alpha=0.6,
                 position=position_dodge(.9)) +
#ylim(-4.75,3)+
    theme_classic()+
    ylab("log2FoldChange")+
    labs(title = "Kegg pathways in human and non-human primate skin microbiomes",fill="") +
  scale_fill_manual(labels = c("Chin", "Forehead"), values = c("#ff6e54","#444e86" ))+
    theme(axis.text.x = element_text(angle = 0,hjust=0.75,vjust=0.2,size=12),
              axis.text.y = element_text(size=12),
          axis.title.y = element_blank(),
          axis.title.x=element_text(size=12),
          title=element_text(size=12))+
   theme(plot.title = element_text(hjust = 0.75,vjust=1))+
    coord_flip()+
 
  theme(legend.text=element_text(size=12))
pK
```
#now doing species

```{r}
cts=read.csv("oguspecies_table.csv",header=TRUE,row.names=1)
cts
meta <- subset(meta, !is.na(Location))
meta=subset(meta,Location!='human')

icol=colnames(cts)
irow=rownames(meta)

int=intersect(irow,icol)
cts2=cts[ ,int]
meta2=meta[int, ]
```
```{r}
cts2[is.na(cts2)] <- 0
dds <- DESeqDataSetFromMatrix(countData = cts2,
                              colData = meta2,
                              design= ~ Location)
                              
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

humanVg=results(dds, contrast=c("Location","Chin","Cheek"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

ChinVCheekSpecies=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(ChinVCheekSpecies), file="deseq2SpeciesChinVCheek.csv")
ChinVCheekSpecies[1:10,]
```
```{r}
pwd=read.csv("deseq2SpeciesChinVCheek.csv",header=TRUE,row.names=1)
pwd2=data.frame(pwd[1:20,])
OTUs=rownames(pwd2)

pK=ggplot(pwd2,aes(reorder(OTUs,-log2FoldChange),log2FoldChange,fill=log2FoldChange < 0))+
    geom_bar(stat="identity")+
    geom_errorbar(aes(ymin=log2FoldChange-lfcSE, ymax=log2FoldChange+(lfcSE)), width=.2, alpha=0.6,
                 position=position_dodge(.9)) +
#ylim(-4.75,3)+
    theme_classic()+
    ylab("log2FoldChange")+
    labs(title = "Differentially abundant microbes between Chin and Cheek microbiomes",fill="") +
  scale_fill_manual(labels = c("Chin", "Cheek"), values = c("#ff6e54","#444e86" ))+
    theme(axis.text.x = element_text(angle = 0,hjust=0.75,vjust=0.2,size=12),
              axis.text.y = element_text(size=12),
          axis.title.y = element_blank(),
          axis.title.x=element_text(size=12),
          title=element_text(size=12))+
   theme(plot.title = element_text(hjust = 0.75,vjust=1))+
    coord_flip()+
 
  theme(legend.text=element_text(size=12))
pK
```

```{r}
humanVg=results(dds, contrast=c("Location","Chin","Forehead"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

ChinVForeheadSpecies=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(ChinVForeheadSpecies), file="deseq2SpeciesChinVForehead.csv")
ChinVForeheadSpecies[1:10,]
```
NONSIG above
now Cheek Forehead
```{r}
humanVg=results(dds, contrast=c("Location","Cheek","Forehead"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

CheekVForeheadSpecies=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(CheekVForeheadSpecies), file="deseq2SpeciesCheekVForehead.csv")
CheekVForeheadSpecies[1:10,]
```

```{r}
```


```{r}
pwd=read.csv("deseq2SpeciesCheekVForehead.csv",header=TRUE,row.names=1)
pwd2=data.frame(pwd[1:20,])
OTUs=rownames(pwd2)

pK=ggplot(pwd2,aes(reorder(OTUs,-log2FoldChange),log2FoldChange,fill=log2FoldChange < 0))+
    geom_bar(stat="identity")+
    geom_errorbar(aes(ymin=log2FoldChange-lfcSE, ymax=log2FoldChange+(lfcSE)), width=.2, alpha=0.6,
                 position=position_dodge(.9)) +
#ylim(-4.75,3)+
    theme_classic()+
    ylab("log2FoldChange")+
    labs(title = "Differentially abundant microbes between Cheek and Forehead microbiomes",fill="") +
  scale_fill_manual(labels = c("Cheek", "Forehead"), values = c("#ff6e54","#444e86" ))+
    theme(axis.text.x = element_text(angle = 0,hjust=0.75,vjust=0.2,size=12),
              axis.text.y = element_text(size=12),
          axis.title.y = element_blank(),
          axis.title.x=element_text(size=12),
          title=element_text(size=12))+
   theme(plot.title = element_text(hjust = 0.75,vjust=1))+
    coord_flip()+
 
  theme(legend.text=element_text(size=12))
pK
```
```{r}
library(data.table)
library(vegan)
library(dplyr)
library(ecodist)
```
```{r}
cts3=t(cts2)

    cts4=bcdist(cts3)
      nrow(cts4)
ad=adonis2(cts4~Location*Species, meta2, permutations = 999, method = "bray")
ad  

adflip=adonis2(cts4~Species+Location, meta, permutations = 999, method = "bray")
adflip
write.csv(file='adonisSpeciesAllChimpandGorilla.csv',ad)
```
```{r}
kts <- read.csv("kegg2koname.csv",header=TRUE,row.names=1)
kts

meta <- subset(meta, !is.na(Location))
meta=subset(meta,Location!='human')

icol=colnames(kts)
irow=rownames(meta)

int=intersect(irow,icol)
kts2=kts[ ,int]
meta2=meta[int, ]
```
```{r}
kts3=t(kts2)

kts4=bcdist(kts3)
      
ad=adonis2(kts4~Location+Species, meta2, permutations = 999, method = "bray")
ad 
```
```{r}


pts <- read.csv("pathway_type.csv",header=TRUE,row.names=1)
pts


icol=colnames(pts)
irow=rownames(meta)

int=intersect(irow,icol)
pts2=pts[ ,int]


pts3=t(pts2)

pts4=bcdist(pts3)
      
ad=adonis2(pts4~Location+Species, meta2, permutations = 999, method = "bray")
ad 
```
Now checking kegg pathway Deseq

```{r}
```{r}
kts2[is.na(kts2)] <- 0
dds <- DESeqDataSetFromMatrix(countData = kts2,
                              colData = meta2,
                              design= ~ Location)
                              
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

humanVg=results(dds, contrast=c("Location","Chin","Cheek"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

ChinVCheekKegg=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(ChinVCheekKegg), file="deseq2KeggChinVCheek.csv")
ChinVCheekKegg[1:10,]
```
```{r}
humanVg=results(dds, contrast=c("Location","Chin","Forehead"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

ChinVForeheadKegg=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(ChinVForeheadKegg), file="deseq2KeggChinVForehead.csv")
ChinVForeheadKegg[1:10,]
```

```{r}
humanVg=results(dds, contrast=c("Location","Cheek","Forehead"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

CheekVForeheadKegg=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(CheekVForeheadKegg), file="deseq2KeggCheekVForehead.csv")
CheekVForeheadKegg[1:10,]
```
```{r}
setwd("/Users/samd/OneDrive\ -\ Northwestern\ University/Primate/Scripts")
library(DESeq2)
library(ggplot2)
```

```{r}
cts=read.csv("oguspecies_table.csv",header=TRUE,row.names=1)
cts
meta <- read.table("Primate_metadata.txt", sep="\t",header=TRUE,row.names=1)
meta
meta <- subset(meta, !is.na(Location))
metaC=subset(meta,Species=='Pan_troglodytes')
metaG=subset(meta,Species=='Gorilla_gorilla_gorilla')



```
```{r}
icol=colnames(cts)
irow=rownames(metaC)

int=intersect(irow,icol)
ctsC=cts[ ,int]
metaC=meta[int, ]

```
```{r}
dds <- DESeqDataSetFromMatrix(countData = ctsC,
                              colData = metaC,
                              design= ~ Location)
                              
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

humanVg=results(dds, contrast=c("Location","Chin","Cheek"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

ChinVCheekC=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(ChinVCheekC), file="deseq2SpeciesChinVCheekC.csv")
```
```{r}
humanVg=results(dds, contrast=c("Location","Chin","Forehead"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

ChinVForeheadC=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(ChinVForeheadC), file="deseq2SpeciesChinVForeheadC.csv")
```

```{r}

pwd2=data.frame(ChinVForeheadC[1:10,])

pwd2=read.csv("deseq2SpeciesChinVForeheadC.csv",header=TRUE, row.names=1)
pwd2=pwd2[1:10,]

OTUs=rownames(pwd2)

pK=ggplot(pwd2,aes(reorder(OTUs,-log2FoldChange),log2FoldChange,fill=log2FoldChange < 0))+
    geom_bar(stat="identity")+
    geom_errorbar(aes(ymin=log2FoldChange-lfcSE, ymax=log2FoldChange+(lfcSE)), width=.2, alpha=0.6,
                 position=position_dodge(.9)) +
#ylim(-4.75,3)+
    theme_classic()+
    ylab("log2FoldChange")+
    labs(title = "Chimpanzee chin vs. forehead",fill="") +
  scale_fill_manual(labels = c("Chin", "Forehead"), values = c("#ff6e54","#444e86" ))+
    theme(axis.text.x = element_text(angle = 0,hjust=0.75,vjust=0.2,size=12),
              axis.text.y = element_text(size=12),
          axis.title.y = element_blank(),
          axis.title.x=element_text(size=12),
          title=element_text(size=12))+
   theme(plot.title = element_text(hjust = 0.75,vjust=1))+
    coord_flip()+
 
  theme(legend.text=element_text(size=12))
pK
```

```{r}
humanVg=results(dds, contrast=c("Location","Cheek","Forehead"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

CheekVForeheadC=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(CheekVForeheadC), file="deseq2SpeciesCheekVForeheadC.csv")
```
Now GOrilla
```{r}
icol=colnames(cts)
irow=rownames(metaG)

int=intersect(irow,icol)
ctsG=cts[ ,int]
metaG=meta[int, ]

```
```{r}
dds <- DESeqDataSetFromMatrix(countData = ctsG,
                              colData = metaG,
                              design= ~ Location)
                              
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

humanVg=results(dds, contrast=c("Location","Chin","Cheek"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

ChinVCheekG=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(ChinVCheekG), file="deseq2SpeciesChinVCheekG.csv")
```
```{r}
pwd2=data.frame(ChinVCheekG[1:5,])
pwd2=read.csv("deseq2SpeciesChinVCheekG.csv",header=TRUE, row.names=1)
pwd2=pwd2[1:10,]
OTUs=rownames(pwd2)

pK=ggplot(pwd2,aes(reorder(OTUs,-log2FoldChange),log2FoldChange,fill=log2FoldChange < 0))+
    geom_bar(stat="identity")+
    geom_errorbar(aes(ymin=log2FoldChange-lfcSE, ymax=log2FoldChange+(lfcSE)), width=.2, alpha=0.6,
                 position=position_dodge(.9)) +
#ylim(-4.75,3)+
    theme_classic()+
    ylab("log2FoldChange")+
    labs(title = "Differentially abundant microbes between Gorilla Chin and Cheek microbiomes",fill="") +
  scale_fill_manual(labels = c("Chin", "Cheek"), values = c("#ff6e54","#444e86" ))+
    theme(axis.text.x = element_text(angle = 0,hjust=0.75,vjust=0.2,size=12),
              axis.text.y = element_text(size=12),
          axis.title.y = element_blank(),
          axis.title.x=element_text(size=12),
          title=element_text(size=12))+
   theme(plot.title = element_text(hjust = 0.75,vjust=1))+
    coord_flip()+
 
  theme(legend.text=element_text(size=12))
pK
```

```{r}
humanVg=results(dds, contrast=c("Location","Chin","Forehead"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

ChinVForeheadG=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(ChinVForeheadG), file="deseq2SpeciesChinVForeheadG.csv")
```
```{r}
humanVg=results(dds, contrast=c("Location","Cheek","Forehead"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

CheekVForeheadG=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(CheekVForeheadG), file="deseq2SpeciesCheekVForeheadG.csv")
```


```{r}
pwd2=data.frame(CheekVForeheadG[1:10,])
pwd2=read.csv("deseq2SpeciesCheekVForeheadG.csv",header=TRUE, row.names=1)
pwd2=pwd2[1:10,]
OTUs=rownames(pwd2)

pK=ggplot(pwd2,aes(reorder(OTUs,-log2FoldChange),log2FoldChange,fill=log2FoldChange < 0))+
    geom_bar(stat="identity")+
    geom_errorbar(aes(ymin=log2FoldChange-lfcSE, ymax=log2FoldChange+(lfcSE)), width=.2, alpha=0.6,
                 position=position_dodge(.9)) +
#ylim(-4.75,3)+
    theme_classic()+
    ylab("log2FoldChange")+
    labs(title = "Gorilla cheek vs. forehead",fill="") +
  scale_fill_manual(labels = c("Cheek", "Forehead"), values = c("#ff6e54","#444e86" ))+
    theme(axis.text.x = element_text(angle = 0,hjust=0.75,vjust=0.2,size=12),
              axis.text.y = element_text(size=12),
          axis.title.y = element_blank(),
          axis.title.x=element_text(size=12),
          title=element_text(size=12))+
   theme(plot.title = element_text(hjust = 0.75,vjust=1))+
    coord_flip()+
 
  theme(legend.text=element_text(size=12))
pK
```
NOW pATHWAYS
first chimp
```{r}
pts <- read.csv("pathway_type.csv",header=TRUE,row.names=1)
pts


icol=colnames(pts)
irow=rownames(metaC)

int=intersect(irow,icol)
ptsC=pts[ ,int]
```
```{r}
dds <- DESeqDataSetFromMatrix(countData = ptsC,
                              colData = metaC,
                              design= ~ Location)
                              
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

humanVg=results(dds, contrast=c("Location","Chin","Cheek"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

ChinVCheekCpathway=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(ChinVCheekCpathway), file="deseq2PathwayChinVCheekC.csv")
```
```{r}
humanVg=results(dds, contrast=c("Location","Chin","Forehead"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

ChinVForeheadCpathway=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(ChinVForeheadCpathway), file="deseq2PathwayChinVForeheadC.csv")
```
```{r}
humanVg=results(dds, contrast=c("Location","Cheek","Forehead"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

CheekVForeheadCpathway=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(CheekVForeheadCpathway), file="deseq2PathwayCheekVForeheadC.csv")
```
Now Gorilla
```{r}
pts <- read.csv("pathway_type.csv",header=TRUE,row.names=1)
pts


icol=colnames(pts)
irow=rownames(metaG)

int=intersect(irow,icol)
ptsG=pts[ ,int]
```
```{r}
dds <- DESeqDataSetFromMatrix(countData = ptsG,
                              colData = metaG,
                              design= ~ Location)
                              
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

humanVg=results(dds, contrast=c("Location","Chin","Cheek"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

ChinVCheekGpathway=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(ChinVCheekGpathway), file="deseq2PathwayChinVCheekG.csv")
```
```{r}
humanVg=results(dds, contrast=c("Location","Chin","Forehead"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

ChinVForeheadGpathway=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(ChinVForeheadGpathway), file="deseq2PathwayChinVForeheadG.csv")
```
```{r}
humanVg=results(dds, contrast=c("Location","Cheek","Forehead"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

CheekVForeheadGpathway=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(CheekVForeheadGpathway), file="deseq2PathwayCheekVForeheadG.csv")
```
~~~~~~~~~~~~~~~
DOing parblot of adonis results for taxonomy and face site

```{r}
#make sure ggplot2 is installed
library(ggplot2)

#read in the table
adskin=read.csv("adonisSpeciesAllChimpandGorilla_modifiedggplot2.csv", header=TRUE)
adskin

#This is to set order of the factor labels (not necessary initially can skip if u want)
adskin$Factor <- factor(adskin$Factor, levels = c("Face site", "Host"))
adskin$Type <- factor(adskin$Type, levels = c("Taxonomy", "Pathways"))

#This is to make the plot
  (p=ggplot(data=adskin, aes(x=Factor, y=R2)) +
  geom_bar(aes(fill=Factor),stat="identity", position = "dodge")+
  facet_wrap(~Type, scales = "free_x", strip.position="bottom")+
  scale_y_continuous(limits = c(0,0.45), expand = c(0, 0)) +
    scale_fill_manual(values=c("#B18742","#333738"))+
    ylab("R-squared")+
      xlab("")+
     theme_classic()+
    theme(text=element_text(size=17))+
  theme(strip.background = element_blank(),strip.placement="outside"))
    
```
Now trying kegg specifically

```{r}
kts <- read.csv("kegg2koname.csv",header=TRUE,row.names=1)
kts

meta <- subset(meta, !is.na(Location))
meta=subset(meta,Location!='human')

#chimp subset
icol=colnames(kts)
irow=rownames(metaC)

int=intersect(irow,icol)
ktsC=kts[ ,int]

#gorilla subset
icol=colnames(kts)
irow=rownames(metaG)

int=intersect(irow,icol)
ktsG=kts[ ,int]

```
```{r}
dds <- DESeqDataSetFromMatrix(countData = ktsC,
                              colData = metaC,
                              design= ~ Location)
                              
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

humanVg=results(dds, contrast=c("Location","Chin","Cheek"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

ChinVCheekCpathway=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(ChinVCheekCpathway), file="deseq2keggChinVCheekC.csv")
```
```{r}
humanVg=results(dds, contrast=c("Location","Chin","Forehead"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

ChinVForeheadCpathway=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(ChinVForeheadCpathway), file="deseq2keggChinVForeheadC.csv")
```
```{r}
humanVg=results(dds, contrast=c("Location","Cheek","Forehead"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

CheekVForeheadCpathway=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(CheekVForeheadCpathway), file="deseq2keggCheekVForeheadC.csv")
```

```{r}
dds <- DESeqDataSetFromMatrix(countData = ktsG,
                              colData = metaG,
                              design= ~ Location)
                              
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

humanVg=results(dds, contrast=c("Location","Chin","Cheek"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

ChinVCheekGpathway=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(ChinVCheekCpathway), file="deseq2keggChinVCheekG.csv")
```
```{r}
humanVg=results(dds, contrast=c("Location","Chin","Forehead"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

ChinVForeheadGpathway=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(ChinVForeheadGpathway), file="deseq2keggChinVForeheadG.csv")
```

```{r}
humanVg=results(dds, contrast=c("Location","Cheek","Forehead"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

CheekVForeheadGpathway=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(CheekVForeheadGpathway), file="deseq2keggCheekVForeheadG.csv")
```
~~~~~~~~~~~~~~~~~~~~~~~
And now redoing human vs non human analyses with new files with more samples

```{r}
setwd("/Users/samd/OneDrive\ -\ Northwestern\ University/Primate/Scripts")
meta=read.table("metadata2_Indiv.txt",sep="\t",header=TRUE,row.names=1)


cts=read.csv("pathway_type_barplot2.csv",header=TRUE,row.names=1)
cts

```

```{r}
#meta <- subset(meta, !is.na(Location))
#save this for face site redo down below: meta=subset(meta,Location!='human')

icol=colnames(cts)
irow=rownames(meta)

int=intersect(irow,icol)
cts2=cts[ ,int]
meta2=meta[int, ]


dds <- DESeqDataSetFromMatrix(countData = cts2,
                              colData = meta2,
                              design= ~ human)
                              
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

humanVg=results(dds, contrast=c("human","yes","no"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

humanVnhppath=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(humanVnhppath), file="deseq2pathHumanVnhp.csv")
```

```{r}
pwd2=read.csv("deseq2pathHumanVnhp.csv",header=TRUE, row.names=1)
pwd2=pwd2[1:10,]
OTUs=rownames(pwd2)

pK=ggplot(pwd2,aes(reorder(OTUs,-log2FoldChange),log2FoldChange,fill=log2FoldChange < 0))+
    geom_bar(stat="identity")+
    geom_errorbar(aes(ymin=log2FoldChange-lfcSE, ymax=log2FoldChange+(lfcSE)), width=.2, alpha=0.6,
                 position=position_dodge(.9)) +
#ylim(-4.75,3)+
    theme_classic()+
    ylab("log2FoldChange")+
    labs(title = "Skin microbial pathways across humans and",fill="") +
  scale_fill_manual(labels = c("Human", "Non-human primate"), values = c("#ff6e54","#444e86" ))+
    theme(axis.text.x = element_text(angle = 0,hjust=0.75,vjust=0.2,size=12),
              axis.text.y = element_text(size=12),
          axis.title.y = element_blank(),
          axis.title.x=element_text(size=12),
          title=element_text(size=12))+
   theme(plot.title = element_text(hjust = 0.75,vjust=1))+
    coord_flip()+
 
  theme(legend.text=element_text(size=12))
pK
```
Now trying to merge because results are biased towards NHP
Merging by SID
```{r}
setwd("/Users/samd/OneDrive\ -\ Northwestern\ University/Primate/Scripts")
metaphy=read.table("metadata2_Indiv.txt",sep="\t",header=TRUE,row.names=1)


ctsphy=read.csv("pathway_type_barplot2.csv",header=TRUE,row.names=1)
ctsphy

OTU = otu_table(ctsphy, taxa_are_rows = TRUE)
META=sample_data(metaphy)
phyp=phyloseq(OTU,META)

phyp2 = merge_samples(phyp, "SID")
phyp3=data.frame(t(phyp2@otu_table))
#had to add Xs cuz of R..not ideal
sampleid=c( "X1" , "X2" , "X3" , "X4" , "X5" , "X6" , "X7" , "DN" ,"JN", "PP", "ST")
Species=c("Gorilla","Chimp","Chimp","Gorilla","Gorilla","Gorilla","Chimp","human","human","human","human")
human=c("no","no","no","no","no","no","no","yes","yes","yes","yes")
mergedmeta=data.frame(sampleid,Species,human)
rownames(mergedmeta)=mergedmeta$sampleid
mergedmeta



dds <- DESeqDataSetFromMatrix(countData = phyp3,
                              colData = mergedmeta,
                              design= ~ human)
                              
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

humanVg=results(dds, contrast=c("human","yes","no"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

humanVnhpMergedpath=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(humanVnhpMergedpath), file="deseq2pathhumanVnhpMerged.csv")

```

```{r}
#pwd2=read.csv("deseq2pathHumanVnhp.csv",header=TRUE, row.names=1)
pwd2=data.frame(humanVnhpMergedpath[1:10,])
OTUs=rownames(pwd2)

pK=ggplot(pwd2,aes(reorder(OTUs,-log2FoldChange),log2FoldChange,fill=log2FoldChange < 0))+
    geom_bar(stat="identity")+
    geom_errorbar(aes(ymin=log2FoldChange-lfcSE, ymax=log2FoldChange+(lfcSE)), width=.2, alpha=0.6,
                 position=position_dodge(.9)) +
#ylim(-4.75,3)+
    theme_classic()+
    ylab("log2FoldChange")+
    labs(title = "Skin microbial pathways across humans and NHPs",fill="") +
  scale_fill_manual(labels = c("Human", "Non-human primate"), values = c("#ff6e54","#444e86" ))+
    theme(axis.text.x = element_text(angle = 0,hjust=0.75,vjust=0.2,size=12),
              axis.text.y = element_text(size=12),
          axis.title.y = element_blank(),
          axis.title.x=element_text(size=12),
          title=element_text(size=12))+
   theme(plot.title = element_text(hjust = 0.75,vjust=1))+
    coord_flip()+
 
  theme(legend.text=element_text(size=12))
pK
```
```{r}
setwd("/Users/samd/OneDrive\ -\ Northwestern\ University/Primate/Scripts")
ntphy=read.csv("pathwaynotype.csv",header=TRUE,row.names=1)
ntphy

OTUn = otu_table(ntphy, taxa_are_rows = TRUE)
METAn=sample_data(metaphy)
phypn=phyloseq(OTUn,METAn)

phyp2n = merge_samples(phypn, "SID")
phyp3n=data.frame(t(phyp2n@otu_table))
#had to add Xs cuz of R..not ideal
sampleid=c( "X1" , "X2" , "X3" , "X4" , "X5" , "X6" , "X7" , "DN" ,"JN", "PP", "ST")
Species=c("Gorilla","Chimp","Chimp","Gorilla","Gorilla","Gorilla","Chimp","human","human","human","human")
human=c("no","no","no","no","no","no","no","yes","yes","yes","yes")
mergedmeta=data.frame(sampleid,Species,human)
rownames(mergedmeta)=mergedmeta$sampleid
mergedmeta



dds <- DESeqDataSetFromMatrix(countData = phyp3n,
                              colData = mergedmeta,
                              design= ~ human)
                              
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

humanVg=results(dds, contrast=c("human","yes","no"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

humanVnhpMergedpathNoType=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(humanVnhpMergedpathNoType), file="deseq2pathhumanVnhpMergedNoType.csv")
```
```{r}
#pwd2=read.csv("deseq2pathHumanVnhp.csv",header=TRUE, row.names=1)
pwd2=data.frame(humanVnhpMergedpathNoType[1:10,])
OTUs=rownames(pwd2)

pK=ggplot(pwd2,aes(reorder(OTUs,-log2FoldChange),log2FoldChange,fill=log2FoldChange < 0))+
    geom_bar(stat="identity")+
    geom_errorbar(aes(ymin=log2FoldChange-lfcSE, ymax=log2FoldChange+(lfcSE)), width=.2, alpha=0.6,
                 position=position_dodge(.9)) +
#ylim(-4.75,3)+
    theme_classic()+
    ylab("log2FoldChange")+
    labs(title = "Skin microbial pathways (MetaCyc) across humans and NHPs",fill="") +
  scale_fill_manual(labels = c("Human", "Non-human primate"), values = c("#ff6e54","#444e86" ))+
    theme(axis.text.x = element_text(angle = 0,hjust=0.75,vjust=0.2,size=12),
              axis.text.y = element_text(size=12),
          axis.title.y = element_blank(),
          axis.title.x=element_text(size=12),
          title=element_text(size=12))+
   theme(plot.title = element_text(hjust = 0.75,vjust=1))+
    coord_flip()+
 
  theme(legend.text=element_text(size=12))
pK
```
Now doing face site

```{r}
ntphy=read.csv("pathwaynotype.csv",header=TRUE,row.names=1)
ntphy
metaphy=read.table("metadata2_Indiv.txt",sep="\t",header=TRUE,row.names=1)
metaphy <- subset(metaphy, !is.na(Location))
metaphy=subset(metaphy,Location!='human')
metaG=subset(metaphy,Species=='Gorilla_gorilla_gorilla')
metaC=subset(metaphy,Species=='Pan_troglodytes')

#Gorilla first
icol=colnames(ntphy)
irow=rownames(metaG)

int=intersect(irow,icol)
cts2=ntphy[ ,int]
meta2=metaG[int, ]


dds <- DESeqDataSetFromMatrix(countData = cts2,
                              colData = meta2,
                              design= ~ Location)
                              
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

humanVg=results(dds, contrast=c("Location","Chin","Cheek"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

gorillaChinVCheekpath=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(gorillaChinVCheekpath), file="deseq2pathgorillaChinVCheekPathnt.csv")
```

NOW doing FACE SITE with 16S
library(plyr)
```{r}
setwd("/Users/samd/OneDrive\ -\ Northwestern\ University/Primate/Scripts")
of=read.csv("16Sr3otu.csv",header=TRUE,row.names=1)
of
tf=read.csv("16Sr3tax.csv",header=TRUE,row.names=1)
tf
#also trying normal taxtable with full taxaonomy concatenated
tft=read.table("16Sr3tax.txt",header=TRUE,row.names=1,sep="\t")
tft

mf2=read.csv("16Smetadata2.csv",header=TRUE,row.names=1)
mf2

of2=merge(y = of, x = tft, by = 'row.names',all.y=TRUE)
of2

of3=subset(of2,Taxon!="d__Bacteria" & Taxon!="Unassigned")
of4=ddply(of3,"Taxon",numcolwise(sum))
of5 <- of4[,-1]
rownames(of5) <- of4[,1]
of5


```

```{r}
mphy=read.table("16Smetadata2.txt",header=TRUE,row.names=1)
mphy
metaphy <- subset(mphy, !is.na(Location))
metaphy=subset(metaphy,Location!='human')
metaG=subset(metaphy,Species=='Gorilla_gorilla_gorilla')
metaC=subset(metaphy,Species=='Pan_troglodytes')

#Gorilla first
icol=colnames(of5)
irow=rownames(metaG)

int=intersect(irow,icol)
cts2=of5[ ,int]
meta2=metaG[int, ]

library(DESeq2)
dds <- DESeqDataSetFromMatrix(countData = cts2,
                              colData = meta2,
                              design= ~ Location)
                              
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

humanVg=results(dds, contrast=c("Location","Chin","Cheek"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

gorillaChinVCheek16S=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(gorillaChinVCheek16S), file="deseq2pathgorillaChinVCheek16S.csv")
```
```{r}
humanVg=results(dds, contrast=c("Location","Chin","Forehead"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

gorillaChinVFore16S=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(gorillaChinVFore16S), file="deseq2pathgorillaChinVFore16S.csv")
```

```{r}
humanVg=results(dds, contrast=c("Location","Cheek","Forehead"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

gorillaCheekVFore16S=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(gorillaCheekVFore16S), file="deseq2pathgorillaCheekVFore16S.csv")
```

```{r}
#Chimp now
icol=colnames(of5)
irow=rownames(metaC)

int=intersect(irow,icol)
cts2=of5[ ,int]
meta2=metaC[int, ]


dds <- DESeqDataSetFromMatrix(countData = cts2,
                              colData = meta2,
                              design= ~ Location)
                              
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

humanVg=results(dds, contrast=c("Location","Chin","Cheek"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

ChimpChinVCheek16S=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(ChimpChinVCheek16S), file="deseq2pathChimpChinVCheek16S.csv")
```
```{r}
humanVg=results(dds, contrast=c("Location","Chin","Forehead"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

ChimpChinVFore16S=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(ChimpChinVFore16S), file="deseq2pathChimpChinVFore16S.csv")
```
```{r}
humanVg=results(dds, contrast=c("Location","Cheek","Forehead"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

ChimpCheekVFore16S=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(ChimpCheekVFore16S), file="deseq2pathChimpCheekVFore16S.csv")
```
Now doing the 16S taxonomy across human and non human

```{r}
library(qiime2R)

setwd("/Users/samd/OneDrive\ -\ Northwestern\ University/Primate/Scripts")
pg=qza_to_phyloseq(
  features = "16Sr3tablef_grouped_species.qza",
  #taxonomy= "16Sr3tax.qza",
  metadata = "16Smetadata2_grouped.txt"
)
pg
```
```{r}

destroyX = function(es) {
  f = es
  for (col in c(1:ncol(f))){ #for each column in dataframe
    if (startsWith(colnames(f)[col], "X") == TRUE)  { #if starts with 'X' ..
      colnames(f)[col] <- substr(colnames(f)[col], 2, 100) #get rid of it
    }
  }
  assign(deparse(substitute(es)), f, inherits = TRUE) #assign corrected data to original name
}

pgm=data.frame(pg@sam_data)
pgo=data.frame(pg@otu_table)
destroyX(pgo)
pgo
```

```{r}
library(dplyr)

pgm1=pgm %>% mutate(HumanYN = ifelse(pgm$Species=="Homo_sapiens", "Human", "NHP"))
pgm1



dds <- DESeqDataSetFromMatrix(countData = pgo,
                              colData = pgm1,
                              design= ~ HumanYN)
                              
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

humanVg=results(dds, contrast=c("HumanYN","Human","NHP"))
res2=humanVg[order(-humanVg$padj), "pvalue", drop = FALSE]

HumanVsNon16S_species=humanVg[order(humanVg$padj),]

write.csv(as.data.frame(HumanVsNon16S_species), file="deseq2HumanVsNon16S_species.csv")
```

```{r}
pwd2=read.csv("deseq2HumanVsNon16S_species.csv",header=TRUE, row.names=1)
#pwd2=data.frame(HumanVsNon16S_species[1:10,])
pwd2=pwd2[1:10,]
OTUs=rownames(pwd2)

pK=ggplot(pwd2,aes(reorder(OTUs,-log2FoldChange),log2FoldChange,fill=log2FoldChange < 0))+
    geom_bar(stat="identity")+
    geom_errorbar(aes(ymin=log2FoldChange-lfcSE, ymax=log2FoldChange+(lfcSE)), width=.2, alpha=0.6,
                 position=position_dodge(.9)) +
#ylim(-4.75,3)+
    theme_classic()+
    ylab("log2FoldChange")+
    labs(title = "Skin microbiome differential abundance across humans and NHPs",fill="") +
  scale_fill_manual(labels = c("Human", "Non-human primate"), values = c("#ff6e54","#444e86" ))+
    theme(axis.text.x = element_text(angle = 0,hjust=0.75,vjust=0.2,size=12),
              axis.text.y = element_text(size=12),
          axis.title.y = element_blank(),
          axis.title.x=element_text(size=12),
          title=element_text(size=12))+
   theme(plot.title = element_text(hjust = 0.75,vjust=1))+
    coord_flip()+
 
  theme(legend.text=element_text(size=12))
pK
```

trying to run dada2 on yamomami

```{r}
library(dada2)
path="/Users/samd/Library/CloudStorage/OneDrive-NorthwesternUniversity/Primate/Scripts/yamomami"
list.files(path)
fnFs <- sort(list.files(path, pattern=".fastq", full.names = TRUE))
plotQualityProfile(fnFs[1:2])
```
doing adonis again

```{r}
icol=colnames(of5)
irow=rownames(mf2)

int=intersect(irow,icol)
adt=of5[ ,int]
mad=mf2[int, ]

adt2=t(adt)

    adt3=bcdist(adt2)

adflip=adonis2(adt3~Species+Location, mad, permutations = 999, method = "bray")
adflip
write.csv(file='adonisSpeciesAllChimpandGorilla2.csv',adflip)

```

redoing deseq for NHp vs human MG
```{r}
pw=read.table("deseq2pathHumanVnhp_filtered.csv",sep=',',row.names=1,header=TRUE)
pw
```
```{r}

#pwd2=data.frame(HumanVsNon16S_species[1:10,])
pwd2=pw[1:14,]
OTUs=rownames(pwd2)

pK=ggplot(pwd2,aes(reorder(OTUs,-log2FoldChange),log2FoldChange,fill=log2FoldChange < 0))+
    geom_bar(stat="identity")+
    geom_errorbar(aes(ymin=log2FoldChange-lfcSE, ymax=log2FoldChange+(lfcSE)), width=.2, alpha=0.6,
                 position=position_dodge(.9)) +
#ylim(-4.75,3)+
    theme_classic()+
    ylab("log2FoldChange")+
    labs(title = "Skin microbiome pathway differential abundance across humans and NHPs",fill="") +
  scale_fill_manual(labels = c("Human", "Non-human primate"), values = c("#ff6e54","#444e86" ))+
    theme(axis.text.x = element_text(angle = 0,hjust=0.75,vjust=0.2,size=12),
              axis.text.y = element_text(size=12),
          axis.title.y = element_blank(),
          axis.title.x=element_text(size=12),
          title=element_text(size=12))+
   theme(plot.title = element_text(hjust = 0.75,vjust=1))+
    coord_flip()+
 
  theme(legend.text=element_text(size=12))
pK
```

